Elastic Scattering in General Relativistic Ray Tracing for Neutrinos

### 2017-07-05 paper.tex
Add momentum conventions appendix and simplify upstream text

Because $\ell_\beta$ is subject to two constraints
(normalization and orthogonality to the observer's velocity)
it has two degrees of freedom; we make this explicit by defining its
spatial components with respect to spherical polar angles
\begin{equation}
  \label{eqn:def_direction}
  \ell_\alpha :=
  q (s,\sin\upsilon_1\cos\upsilon_2,\sin\upsilon_1\sin\upsilon_2,\cos\upsilon_1),
\end{equation}
with $q$ and $s$ functions of $\upsilon_1$ and $\upsilon_2$.
Now our symbol for the distribution function,
$f(x^\alpha;\varepsilon,\upsilon_1,\upsilon_2)$,
makes manifest its seven independent arguments.
\todo{also decompose wrt $p_t$}

\subsection{Boltzmann Equation}
\label{ssec:boltzmann}
Neutrino radiation obeys the relativistic Boltzmann Equation,
which, suppressing the arguments of $f$ for simplicity, is written,
\todo{point out limit of QKEs}
\begin{equation}
  \label{eqn:boltzmann}
  \frac{d}{d\lambda}f = C[f],
\end{equation}
where $d/d\lambda$ denotes a derivative with respect to the affine
parameter defining the neutrino momentum (Eqn.~\ref{eqn:geodesic_x})
and $C[f]$ is the source term arising from interactions with the medium,
having dimension ${\rm energy}\,{\rm length}^{-1}$.
\todo{describe $C[f]$ in physical terms}
The source term varies over phase space $(x^\alpha,p_\beta)$,
and due to Fermi blocking depends locally on the distribution function $f$
and nonlocally on the distribution function of this neutrino and its
antiparticle at different momenta, $f'$ and $\bar{f}'$.
For simplicity we symbolize all of these dependencies with the shorthand $C[f]$.
The various neutrino interactions contributing to $C[f]$ are detailed in
App.~\ref{sec:source_terms}.

@@

\subsection{Moments of the Distribution Function}
\label{ssec:moments}
\todo{code these up}
We may take angular moments of the distribution function:
\begin{align}
  \label{eqn:J}
  J(\varepsilon) &=
  \frac{\varepsilon^3}{(2\pi)^3} \oint d\Omega' f(\varepsilon, \ell'_\beta) \\
  \label{eqn:Ha}
  H^\mu(\varepsilon) &=
  \frac{\varepsilon^3}{(2\pi)^3} \oint d\Omega' f(\varepsilon, \ell'_\beta) \ell'^\mu \\
  \label{eqn:Sab}
  S^{\mu\gamma}(\varepsilon) &=
  \frac{\varepsilon^3}{(2\pi)^3} \oint d\Omega' f(\varepsilon, \ell'_\beta) \ell'^\mu \ell'^\gamma,
\end{align}
representing specific energy density, specific energy flux, and
specific momentum flux, respectively.
Here ``specific'' refers to the quantity being integrable over neutrino energy.
\todo{relate moments to other frames?}
Integrals~\ref{eqn:J}--\ref{eqn:Sab} are performed over a solid angle in
momentum space while holding $\varepsilon$ constant:
$d\Omega \equiv d(\cos\upsilon_1)\,d\upsilon_2$.
\todo{consistent with Eqn.~\ref{eqn:def_direction}?}
We will also make use of the number density and flux defined as
$G(\varepsilon)=J(\varepsilon)/\varepsilon$
and $K^\mu(\varepsilon)=H^\mu(\varepsilon)/\varepsilon$
respectively.

@@

We distinguish moments computed assuming an Eulerian observer with a tilde,
e.g. $\tilde{J}$, $\tilde{H}^\mu$, $\tilde{S}^{\mu\nu}$.
Note that these three Eulerian moments are identical to the lab-frame moments
$E$, $F^\mu$, and $P^{\mu\nu}$
defined in
\cite{shib2011-truncated_moment, ocon2015-gr1d_with_nu, fouc2015-m1_nsbh}.

@@

>>>>

Because $\ell_\beta$ is subject to two constraints
(normalization and orthogonality to the observer's velocity)
it has two degrees of freedom; we make this explicit by defining its
spatial cartesian components with respect to spherical polar angles
\begin{equation}
  \label{eqn:def_direction}
  \ell_\alpha \rightarrow
  q (s,\sin a \cos b,\sin a\sin b,\cos a),
\end{equation}
with $q$ and $s$ functions of $a$ and $b$.
Now our symbol for the distribution function,
$f(x^\alpha;\varepsilon,a,b)$,
makes manifest its seven independent arguments.

We also define a rotated frame
\begin{equation}
  \ell_{\alpha'}=\frac{\partial x^\alpha}{\partial x^{\alpha'}}\,\ell_\alpha \nonumber
\end{equation}
in which the cartesian components of the direction 1-form are defined
\begin{equation}
  \label{eqn:def_direction_primed}
  \ell_{\alpha'} \rightarrow
  q (s,\sin A \cos B,\sin A\sin B,\cos A),
\end{equation}
so that momenta with vanishing polar angle $A$ move outward along coordinate
radial rays.
This transformation is chosen so that for an observer far from the source,
incoming radiation will be concentrated into a narrow beam around $\cos A\approx1$,
independent of that observer's position in coordinate space.
For explicit definitions see App.~\ref{sec:def_momentum}.

In Minkowski spacetime, and for a stationary observer
$u_\beta\rightarrow(-1,0,0,0)$, we would have $s=0$, $q=1$,
and $\varepsilon=-p_t$. In that case $\cos A$ may be identified
\todo{cite someone that uses $\mu$}
with the familiar forward direction cosine $\mu$.

\subsection{Boltzmann Equation}
\label{ssec:boltzmann}
Neutrino radiation obeys the relativistic Boltzmann Equation,
which, suppressing the arguments of $f$ for simplicity, is written,
\todo{point out limit of QKEs}
\begin{equation}
  \label{eqn:boltzmann}
  \frac{d}{d\lambda}f = C[f],
\end{equation}
where $d/d\lambda$ denotes a derivative with respect to the affine
parameter defining the neutrino momentum (Eqn.~\ref{eqn:geodesic_x})
and $C[f]$ is the source term arising from interactions with the medium,
having dimension ${\rm energy}\,{\rm length}^{-1}$.
The source term varies over phase space $(x^\alpha,p_\beta)$,
and due to Fermi blocking depends locally on the distribution function $f$
and nonlocally on the distribution function of this neutrino and its
antiparticle at different momenta, $f'$ and $\bar{f}'$.
For simplicity we symbolize all of these dependencies with the shorthand $C[f]$.
The various neutrino interactions contributing to $C[f]$ are detailed in
App.~\ref{sec:source_terms}.

@@

\subsection{Moments of the Distribution Function}
\label{ssec:moments}
\todo{code these up}
We may take angular moments of the distribution function:
\begin{align}
  \label{eqn:J}
  J(\varepsilon) &=
  \frac{\varepsilon^3}{(2\pi)^3} \oint d\Omega' f(\varepsilon, \ell'_\beta) \\
  \label{eqn:Ha}
  H^\mu(\varepsilon) &=
  \frac{\varepsilon^3}{(2\pi)^3} \oint d\Omega' f(\varepsilon, \ell'_\beta) \ell'^\mu \\
  \label{eqn:Sab}
  S^{\mu\gamma}(\varepsilon) &=
  \frac{\varepsilon^3}{(2\pi)^3} \oint d\Omega' f(\varepsilon, \ell'_\beta) \ell'^\mu \ell'^\gamma,
\end{align}
representing specific energy density, specific energy flux, and
specific momentum flux, respectively.
Here ``specific'' refers to the quantity being integrable over neutrino energy.
\todo{relate moments to other frames?}
Integrals~\ref{eqn:J}--\ref{eqn:Sab} are performed over a solid angle in
momentum space while holding $\varepsilon$ constant:
$d\Omega \equiv d(\cos a)\,db$.
\todo{consistent with Eqn.~\ref{eqn:def_direction}?}
We will also make use of the number density and flux defined as
$G(\varepsilon)=J(\varepsilon)/\varepsilon$
and $K^\mu(\varepsilon)=H^\mu(\varepsilon)/\varepsilon$
respectively.

@@

There are two sensible observers to use to define the frame in which these
moments are computed, that is defining the four-velocity of
Eqn.~\ref{eqn:def_momentum}:
a stationary (also called an Eulerian) observer,
and a fluid (also called a comoving) observer \cite{smar1980-gr_hydro}.
Definitions are given in App.~\ref{sec:def_momentum}.
We distinguish moments computed assuming an Eulerian observer with a tilde,
e.g. $\tilde{J}$, $\tilde{H}^\mu$, $\tilde{S}^{\mu\nu}$.
Note that these three Eulerian moments are identical to the lab-frame moments
$E$, $F^\mu$, and $P^{\mu\nu}$
defined in
\cite{shib2011-truncated_moment, ocon2015-gr1d_with_nu, fouc2015-m1_nsbh}.

@@

\section{Momentum Definitions}
\label{sec:def_momentum}
Here we describe our conventions defining the four-momentum of a neutrino.
We decompose the neutrino momentum into components parallel and orthogonal to
an observer's velocity $u_\beta$:
\begin{equation}
  \label{eqn:def_momentum_2}
  p_\beta = \varepsilon (u_\beta + \ell_\beta),
\end{equation}
with $u_\alpha\ell^\alpha=0$ and $\ell_\alpha\ell^\alpha=1$.

We specify the neutrino direction in the observer's frame
with two spherical polar angles
\begin{align}
  \ell_i &= q\Omega_i, \\
  \Omega_i &:= (\sin a \cos b,\sin a\sin b,\cos a).
\end{align}
with q and s given by
\begin{align}
  q &= W
\end{align}
And we further specify a rotated frame using two new spherical polar angles
\begin{equation}
  \label{eqn:def_direction_primed_2}
  \Omega_{i'} :=  (\sin A \cos B,\sin A\sin B,\cos A).
\end{equation}
The two frames are related by a standard Euler rotation matrix
\begin{equation}
  \label{eqn:rotation_connection}
  \Omega_i=R^{i'}_i(\theta,\phi)\, \ell_{i'}.
\end{equation}
